---
title: "Customer Churn Analysis"
author: "Alphonzo Smith"
date: "5/3/2021"
output: 
    html_document:
        toc: true
        toc_float: true
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(caret)
library(patchwork)
```


# Data Import and Summary 
```{r echo=TRUE}
raw <- read_csv("C:/R/win-library/R-Practice/Data/churn_customer.csv",
  col_types = cols(CustomerSince = col_datetime("%m/%d/%y %I:%M %p")))
str(raw)
```

```{r echo = TRUE}
summary(raw)
```


```{r echo = TRUE}
raw %>% is.na %>% colMeans
```

# Scaler
```{r echo = true}

scaler <- readRDS("C:/Users/alsjr/OneDrive/Documents/College Classes/Info 400/Final Documents/customer-churn-project/rds/churn_scaler.rds")

print(scaler)

```

#Load Model and Scaler to make Predictions 
```{r echo = true}

my_model <- readRDS("C:/Users/alsjr/OneDrive/Documents/College Classes/Info 400/Final Documents/customer-churn-project/rds/linear_model.rds")

scaler <- readRDS("C:/Users/alsjr/OneDrive/Documents/College Classes/Info 400/Final Documents/customer-churn-project/rds/churn_scaler.rds")

new_data <- tibble(Technology4G=c(1, 0),
            Technologyfiber=c(0, 0),
            Technologylandline=c(0, 1),
            Technologyphone=c(0, 0),
            Age=c(35, 65),
            SupportCallsLastYear=c(2, 10),
            AverageBill=c(70, 40),
            duration=c(300, 1200))

# Scale Data 
X <- predict(scaler, new_data)

#Predict
predict(my_model, X)

```



# Data Binning of the Churn Indicator

```{r echo = TRUE}

summary(raw$ChurnIndicator)

```


```{r echo =TRUE}

# 1 2 3 4 5 6,  [1,3) 1 close 3 open, [3],  (3, 6]

temp <- hist(raw$ChurnIndicator,
     breaks = c(0.0,0.1, 1.204),
     right= TRUE,
     freq=TRUE,
     col="burlywood",
     main="ChurnIndicator",
     xlab="", ylab="frequency")
```

```{r echo = TRUE}
temp$counts
```

# Creating a Nominal Variable and adding it to a New Dataframe 

```{r echo = TRUE}

map <- c("0", "1")
breaks <- c(0, 0.1, 1.204)

x <- raw$ChurnIndicator
y_ch <- ifelse(x <= breaks[3], map[1], map[2])

churn_cat <- cut(x,
                 breaks = breaks,
                 include.lowest = TRUE,
                 right = FALSE,
                 labels = map,
                 ordered_result = FALSE)

```



```{r echo = TRUE}

churn <- raw %>% add_column(churn_cat = churn_cat)

str(churn)


```


# Plotting New DataFrame colored by churn_cat of ChurnIndicator

```{r echo = TRUE}

churn %>% 
  ggplot(aes(x=ChurnIndicator)) +
  geom_bar(aes(fill=churn_cat, na.rm = TRUE))+
  scale_y_log10() +
  ggtitle("Frequency in log10")+
  labs(y="")

```

# Feature Engineering 

```{r echo = TRUE}

churn <- churn %>% mutate(duration = 
                 as.numeric(difftime(ymd_hms("2016-01-01 00:00:00"), CustomerSince, units="days"))
               )
summary(churn$duration)

```

```{r echo = TRUE}

max(churn$duration)

min(churn$duration)

```


# Feature Scaling

```{r echo=TRUE}

library(caret)
num_attr <- churn %>%
  select(where(is.numeric)) %>% 
  select(-ChurnIndicator) 

scaler <- preProcess(num_attr, method=c("range"))
num_attr_scale <- predict(scaler, num_attr)

```

```{r echo=TRUE}

churn_norm <- tibble(Technology=churn$Technology, 
                     num_attr_scale,
                     ChurnIndicator=churn$ChurnIndicator,
                     churn_cat=churn$churn_cat)
str(churn_norm)

```

# Exploratory Data Analysis

```{r echo=TRUE}

group <- churn %>% 
  select(where(is.numeric)) %>%
  pivot_longer(everything())

group %>% ggplot() +
  geom_histogram(mapping =aes(x = value), color='white', fill='seagreen', alpha=0.6) +
  facet_wrap(vars(name), scales = "free")+
  labs(x="", y="")


```

# Relationship Between Loyalty and Number of Calls

```{r echo=TRUE}

churn_norm %>% ggplot(aes(x=SupportCallsLastYear, y=ChurnIndicator)) +
  geom_point(aes(color=Technology, size=0.1, alpha=0.1)) +
  geom_jitter(aes(color=Technology, alpha=0.1)) +
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(x="",y="") +
  ggtitle("ChurnIndicator (y axis) vs. SupportCallsLastYear (x axis)")

```

# Relationship Between Loyalty and Age

```{r echo=TRUE}

churn_norm %>% ggplot(aes(x=Age, y=ChurnIndicator)) +
  geom_point(aes(color=Technology, size = 0.1, alpha=0.1)) +
  geom_jitter(aes(color=Technology, alpha=0.1)) +
  theme_minimal() +
  theme(legend.position="bottom")
  labs(x="Age", y="ChurnIndicator")

```

# Relationship between Loyalty and Average Bill

```{r echo=TRUE}

churn_norm %>% ggplot(aes(x=AverageBill, y= ChurnIndicator)) +
  geom_point(aes(color = Technology, size = 0.1, alpha=0.1)) +
  geom_jitter(aes(color=Technology, alpha=0.1)) +
  theme_minimal() +
  theme(legend.position="bottom") +
  ggtitle("ChurnIndicator (y) vs. AverageBill (x)") +
  xlab("") + ylab("")

```


# Relationship Between Loyalty and Technology

```{r echo=TRUE}

churn_norm %>% 
  ggplot(aes(x=Technology, y=ChurnIndicator)) +
    geom_violin(aes(fill=Technology), trim=TRUE) +
    xlab("") + ylab("") +
    theme(legend.position="none") +
    guides(fill=FALSE) +
    ggtitle("ChurnIndicator across Technology groups") +
    theme_classic() +
    coord_flip()

```
# Correlation Analysis 

## Calculate Correlation Coefficients
```{r echo = true}

library(polycor)
corr <- hetcor(churn_norm)

corr$correlations

```


#Visualization of Correlation in both direction and strength
```{r echo = true}

library(corrplot)
corrplot(corr$correlations, 
         order = "hclust", 
         addrect = 2, 
         col = c("white", "black"), 
         bg = "gold2")

```

#Rank the Variables by Correlation
```{r echo = true}

corrmat <- as_tibble(corr$correlations)
corrmat <- corrmat %>% mutate(attribute=colnames(corrmat)) 
rank <- corrmat %>%
  select(attribute, ChurnIndicator) %>%
  mutate(abs.correlation=abs(ChurnIndicator)) %>%
  arrange(desc(abs.correlation))

# Print the rank
rank

```

```{r echo = true}

rank$attribute

```

```{r echo = true}

rank %>% ggplot(aes(x=fct_reorder(attribute,abs.correlation), y=abs.correlation)) +
  geom_point(aes(color=ChurnIndicator), alpha=0.5, size=4) +
  xlab("") + ylab("abs(correlation)") +
  coord_flip() +
  theme_minimal()

```

# One hot encoding
```{r echo = true}

library(caret)
dummifier <- dummyVars(" ~ Technology", data=churn_norm)
dummy <- as_tibble(predict(dummifier, churn_norm))

```

```{r echo = true}

churn_norm <- churn_norm %>% 
  add_column(dummy, .after ="Technology") 
str(churn_norm)  

```

# Reranking after dummify
```{r echo = true}

library("polycor")
corr <- hetcor(churn_norm)
corrmat <- as_tibble(corr$correlations)
corrmat <- corrmat %>% 
  mutate(attribute=colnames(corrmat)) 
rank <- corrmat %>%
  select(attribute, ChurnIndicator) %>%
  mutate(abs.correlation=abs(ChurnIndicator)) %>%
  arrange(desc(abs.correlation))

# Print the rank
rank

```

```{r echo = true}

rank %>% 
  ggplot(aes(x=fct_reorder(attribute,abs.correlation), y=abs.correlation)) +
  geom_point(aes(color=ChurnIndicator), alpha=0.5, size=4) +
  xlab("") + ylab("abs(correlation)") +
  coord_flip() +
  theme_minimal()

```

#Model Fitting

## Splitting data into two subsets
```{r echo = true}

library(caret)
trainIndex <- createDataPartition(churn_norm$ChurnIndicator, 
                                  p = .70, 
                                  list = FALSE, 
                                  times = 1)

## Split the data using the defined partition
train_set <- churn_norm[trainIndex,]
X_train <- train_set %>% select(-c(Technology, churn_cat, ChurnIndicator))
y_train <- train_set$ChurnIndicator

test_set <- churn_norm[-trainIndex,]
X_test <- test_set %>% select(-c(Technology, churn_cat, ChurnIndicator))
y_test <- test_set$ChurnIndicator

```

## caret::train
```{r echo = true}

#lm

#base <- lm(ChurnIndicator ~. -Technology-churn_cat, train_set)
linear <- train(y = y_train,
                x = X_train,
                method = 'lm',
                metric =  "Rsquared"
              )

```

```{r echo = true}

#ridge

library(glmnet)
parameters <- c(seq(0.1, 2, 0.1), seq(2, 5, 0.5), seq(5, 25, 1))
ridge <- train(y = y_train,
               x = X_train,
               method = 'glmnet',
               tuneGrid = expand.grid(alpha = 0, lambda = parameters),
               metric = "Rsquared")

```

#Model Testing
```{r echo = true}

truth <- y_test
prediction1 <- predict(linear, X_test) 
prediction2 <- predict(ridge, X_test) 

library(patchwork)
p1 = ggplot()+
  geom_point(aes(x=seq(1:length(truth)), y=truth), color="tomato", alpha=0.3) +
  geom_point(aes(x=seq(1:length(prediction1)), y=prediction1), color="burlywood", alpha=0.3) +
  ylab("") + xlab("") + theme_classic() +
  ggtitle("linear & lasso")

p2 = ggplot()+
  geom_point(aes(x=seq(1:length(truth)), y=truth), color="tomato", alpha=0.3) +
  geom_point(aes(x=seq(1:length(prediction2)), y=prediction2), color="seagreen", alpha=0.3) +
  ylab("") + xlab("") + theme_classic() 

p1 / p2

```

#R Squared
```{r echo = true}

c(R2(prediction1, y_test), R2(prediction2, y_test))

```

#Print the Models
```{r echo = true}

vars <- rownames(as.matrix(coef(ridge$finalModel, ridge$finalModel$lambdaOpt)))
ridge_weight <-  as.matrix(coef(ridge$finalModel, ridge$finalModel$lambdaOpt))["1"]
linear_weight <- linear$finalModel$coefficients

tibble(vars, ridge_weight, linear_weight) 

```

#Model Deployment 
```{r echo = true}

X_raw <- tibble(Technology4G=c(1),
            Technologyfiber=c(0),
            Technologylandline=c(0),
            Technologyphone=c(0),
            Age=c(35),
            SupportCallsLastYear=c(2),
            AverageBill=c(70),
            duration=c(300))
# Scale data
X <- predict(scaler, X_raw)
# Predict
c(predict(linear, X), predict(ridge, X))

```

#Variable Importance 
```{r}

mod2 <- train(ChurnIndicator ~. -Technology-churn_cat,
               data = churn_norm,
               method = "lm")

```

```{r echo = true}

ggplot(varImp(mod2)) + theme_classic()

```

